Revisiting Multi-Domain Machine Translation

### 2020-04-29 2020/coling2020/multidomain.tex
complete table 3,4,6; missing WDCNMT scores in table 5

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{blue}{#1}}}
\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{blue}{#1}}}

@@

We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc. Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$. $h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but can differ from one domain to the other.

@@

\begin{table*}
  \centering% \small
  \begin{tabular}{|p{1.8cm}|*{10}{r|}} \hline
    % &&&&&& \\
    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & \multicolumn{2}{c|}{Wrong} \\ % \hline
     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{old} &  \multicolumn{1}{c|}{new}\\ \hline
    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5}& -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& 32.55 & \SW{-3.28}\\
    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & 35.92 & \SW{-1.7}\\
    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3} $^+$ & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & 34.91 &\SW{-1.8}\\
    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& 37.01 & \SW{-1.4}\\ 
    \system{TTM} (!)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & & -0.1\\
%    \system{MDL Gated} & & & & & & & & &  \\
    \system{DM} (!)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & &-0.9\\
    \system{ADM} (!)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & &-0.2\\
    \system{WDCMT} (!)     &-0.9    & -0.9  & -1.1 & -1.1  & -2.2 & -2.2  & -0.45 & -0.85 & *** & 0.19 \\
    \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & 31.78 & \SW{-1.3}\\
    \hline
  \end{tabular}
  \caption{Translation performance with variable domain definitions. For the Split experiments, we only report the loss in performance for the concerned test set (\domain{med} or \domain{law}), (*) denote significant losses when domains are split or merged; (+) denotes a significant improvement over \system{Full FT}; \fyTodo{Check significance}\fyTodo{How about wrong ?}}
  \label{tab:redomains}
\end{table*}

@@

     \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & & 50.1\\[-2pt]
%                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
                   & -0.2 & -0.1  & -0.1 & -0.6 & -0.9 & +0.3 & \SW{-0.8} & -0.2 & -0.3\\

    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & & 49.9\\[-2pt]
%                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
                   & +0.3  & -0.1   & -0.6    & \SW{-1.1}  &  \SW{-1.3}  &\SW{-3.5}  & \SW{-1.4}  & -0.1 & -1.2\\

    \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & & 46.3 \\[-2pt]
                    & -0.65 &	0.46 &	-0.44 &	-0.58 &	0.49 &	3.8 &	-0.88 &	0.3 &	-0.4 \\

    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & & 49.4\\[-2pt]
%                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
                    & -0.3  & -0.3  & -0.5    & \SW{-1.1}  &  -1.0  & \SW{-4.0}  & \SW{-1.4}& -0.5 &-1.2\\

    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & & 45.2\\[-2pt]
%                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
                   &+0.6&+0.4&+0.6&-0.1&0.5&0.0&\SW{-1.2}& +0.5 & +0.3\\

    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & &4.4 \\[-2pt]
%                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
                   & +0.3 & -0.8 & -0.8 & -0.2 & -0.4 & \SW{-5.0} & \SW{-1.9} & -0.2 & -1.1\\

    \system{WDCMT}     &34.8  &51.8&47.8&30.7&37.3&49.3&25.9 & &\\[-2pt]
                    &&&&&&&& \\

    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & & 51.8 \\[-2pt]
%                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
                        & -0.2 & +0.5 & 0.0 & -0.4  &  -0.2 & -0.2 & \SW{-0.8} & -0.2 & -0.1 \\
     \hline
  \end{tabular}
  \caption{Ability to handle new a domain: BLEU performance in ``resume training'' mode compared to a complete training session. Averages only take into account the six initial domains (\domain{News} excluded).}
  \label{tab:warmrestart}
  \fyTodo{Fill the table, significancy testing against initial learning condition, for each domain}
  \fyTodo{Need correct number / differences for Mixed-Nat}
  \fyDone{averages include one more domain ? No for comparison}
\end{table*}

@@

\begin{table*}
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{w\domain{avg}} & \multicolumn{1}{c||}{\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
  %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
  %     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 49.0 & 41.6 \\%    & 21.8 \\
    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 48.3 & 40.7  \\% & \SW{21.7} \\
    \system{LDC-Tag}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 48.3 & 39.1         \\% & 22.1 \\ 
    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 48.3 & 40.7   \\% &  23.4 \\
    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 42.8 & 37.8 \\ % & 22.6\\
    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 47.1 & 39.5\\% & 23.3 \\
    \system{WDCMT}       & 36.0 & 53.8 & 48.5 & 31.0 & 39.0 & 63.5 & 45.3 & 39.1\\ % & 20.4 \\
    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 50.7& 41.2\\%  & \SW{21.2} \\
     \hline
  \end{tabular}
  \caption{Translation performance with automatic domains, computed with the original test sets.}
  \label{tab:subdomains}
  \fyDone{Fill the table with correct results,}
\end{table*}

Table~\ref{tab:subdomains} reports the effect of using automatic domain tags on the 6 domain-specific test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system; aggregate numbers were then computed, as well as comparisons with the 6-domain scenarios. The results are very clear and confirm previous observations: even though some clusters are very close, the net effect is a decrease in performance for almost all systems and conditions. In this setting, the best MDMT in our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.\fyTodo{Note about scaling?}\fyTodo{More comparison?}

>>>>

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.
\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}
}\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\jcTodo}[1]{\Todo[JC:]{\textcolor{blue}{#1}}}
\newcommand{\jcDone}[1]{\done[JC]\Todo[JC:]{\textcolor{blue}{#1}}}

@@

We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \cite{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for source sentences; each domain corresponds to a specific source of data, and differs from the other data sources in terms of textual genre, thematic content \cite{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc. Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$. $h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$, but can differ from one domain to the other.

@@

\begin{table*}
  \centering% \small
  \begin{tabular}{|p{1.8cm}|*{10}{r|}} \hline
    % &&&&&& \\
    \hfill Set-up & \multicolumn{2}{c|}{Split} &  \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Split} & \multicolumn{2}{c|}{Merge} & \multicolumn{2}{c|}{Wrong} \\ % \hline
     Model \hfill & \multicolumn{2}{c|}{\domain{med} \footnotesize{(0.5 / 0.5)}} &  \multicolumn{2}{c|}{\domain{med} {\footnotesize (0.25 / 0.75)}} & \multicolumn{2}{c|}{\domain{law} {\footnotesize (0.5 / 0.5)}} & \multicolumn{2}{c|}{\domain{bank}+\domain{law}} &  \multicolumn{1}{c|}{old} &  \multicolumn{1}{c|}{new}\\ \hline
    & \domain{med}$_1$ & \domain{med}$_2$ & \domain{med}$_1$ & \domain{med}$_1$ &  \domain{bank}$_1$ & \domain{bank}$_2$ & \domain{bank} & \domain{law}   & \domain{all} & \domain{News} \\
    \system{FT-Full}      & -0.1 & -0.6 & \SW{-1.5}& -0.2& \SW{-2.3} & \SW{-5.1} &\SW{-1.6} & \SW{-1.4}& 32.55 & \SW{-3.28}\\
    \system{DC-Tag}     & -0.2 & -0.3& +\SB{0.1}$^+$& +0.2& -2.2 & -2.9 & -0.5 & -0.4 & 35.92 & \SW{-1.7}\\
    \system{DC-Feat}    & -0.5 & 0.0 & +\SB{0.3} $^+$ & +0.3 & +0.3 & +0.3 & 0.3 & +0.1 & 34.91 &\SW{-1.8}\\
    \system{LDC-Tag}   & +0.1 & +0.1 & +0.4 & +0.4 & 0.0 &  0.0 &  0.0 & +0.1& 37.01 & \SW{-1.4}\\ 
    \system{TTM} (!)        & -0.2 &  -0.2 & -0.2 & -0.2 & -0.3 &-0.3 &  0.0 & -0.3 & & -0.1\\
%    \system{MDL Gated} & & & & & & & & &  \\
    \system{DM} (!)           & -0.3   & -0.3  & +0.4 & +0.4 & +0.3 & +0.3 & +0.9 & +0.1 & &-0.9\\
    \system{ADM} (!)        & +0.6   & +0.6 & +0.4 & +0.4 & +0.4 & +0.4 &  +0.1 & -0.4 & &-0.2\\
    \system{WDCMT} (!)     & -0.03    & -0.03  & +0.2 & +0.2  & +0.83 & +0.83  & -0.45 & -0.85 & *** & 0.19 \\
    \system{MDL Res}   & -0.2   & -0.07 & +\SB{0.18}$^+$ &+0.0 & -0.89 & -0.89 & +\SB{0.7} & -0.35 & 31.78 & \SW{-1.3}\\
    \hline
  \end{tabular}
  \caption{Translation performance with variable domain definitions. For the Split experiments, we only report the loss in performance for the concerned test set (\domain{med} or \domain{law}), (*) denote significant losses when domains are split or merged; (+) denotes a significant improvement over \system{Full FT}; \fyTodo{Check significance}\fyTodo{How about wrong ?}}
  \label{tab:redomains}
\end{table*}

@@

     \system{DC-Feat}     & 37.4 & 54.9   & 50.0    & 34.7  &  43.9  & 79.6 & 28.9  & 41.2 & 50.1\\[-2pt]
%                   & 37.2 & 54.7   & 49.9    & 34.2  &  43.0  & 79.9 & \SW{28.1}  & &\\
                   & -0.2 & -0.1  & -0.1 & -0.6 & -0.9 & +0.3 & \SW{-0.8} & -0.2 & -0.3\\

    \system{DC-Tag}      &  37.7 & 54.5   & 49.9    &  34.8 &  43.9  & 78.8 & 29.5  & 41.4 & 49.9\\[-2pt]
%                                   & 38.0  & 54.4   & 49.2    & \SW{33.6}  &  \SW{42.7}  &\SW{75.3}  & \SW{28.1}  & & 48.9\\
                   & +0.3  & -0.1   & -0.6    & \SW{-1.1}  &  \SW{-1.3}  &\SW{-3.5}  & \SW{-1.4}  & -0.1 & -1.2\\

    \system{LDC-Tag}    & 37.0   & 54.6  & 49.6    & 34.3  &  43.0  &77.0  & 28.7 & 40.8 & 49.2 \\[-2pt]
                    & -0.65 &	+0.46 &	-0.44 &	-0.58 &	+0.49 &	+\SB{3.8} &	-0.88 &	+0.5 &	-0.4 \\

    \system{TTM}           &  37.3 & 54.4   & 49.6    & 33.8  &  42.9  &78.2  & 29.1 & 41 & 49.4\\[-2pt]
%                   & 37.1  & 54.2   & 49.1    & \SW{32.8}  &  42.0  &\SW{74.2}  & \SW{27.7}& 48.2\\
                    & -0.3  & -0.3  & -0.5    & \SW{-1.1}  &  \SW{-1.0}  & \SW{-4.0}  & \SW{-1.4}& -0.5 & -1.2\\

    \system{DM}            &36.0 &51.3&46.8&31.8&39.8&65.7&27.0 & 38.9 & 45.2\\[-2pt]
%                   &36.5&51.7&47.4&31.7&40.3&65.7&\SW{25.8}& & \\
                   &+0.6&+0.4&+0.6&-0.1&0.5&0.0&\SW{-1.2}& +0.5 & +0.3\\

    \system{ADM}          &36.6&54.2&49.1&32.9&42.1&75.7&28.7 & 40.2 & 48.4 \\[-2pt]
%                   &  36.9 &53.5&48.3&32.7&41.7&\SW{70.7}&\SW{26.8}& \\
                   & +0.3 & -0.8 & -0.8 & -0.2 & -0.4 & \SW{-5.0} & \SW{-1.9} & -0.2 & -1.1\\

    \system{WDCMT}     &&&&&&& & &\\[-2pt]
                    &&&&&&&&& \\

    \system{MDL-Res}    &37.7   & 55.6   & 51.1   & 34.4  & 44.5  & 87.5  & 29.1 & 41.9 & 51.8 \\[-2pt]
%                      & 37.5 & 56.1 & 51.1 & 34.0  &  44.3 & 87.3 & \SW{28.4} & & 51.7 & \\
                        & -0.2 & +0.5 & 0.0 & -0.4  &  -0.2 & -0.2 & \SW{-0.8} & -0.2 & -0.1 \\
     \hline
  \end{tabular}
  \caption{Ability to handle new a domain: BLEU performance in ``resume training'' mode compared to a complete training session. Averages only take into account the six initial domains (\domain{News} excluded).}
  \label{tab:warmrestart}
  \fyTodo{Fill the table, significancy testing against initial learning condition, for each domain}
  \fyTodo{Need correct number / differences for Mixed-Nat}
  \fyDone{averages include one more domain ? No for comparison}
\end{table*}

@@

\begin{table*}
  \centering
  \fyDone{Fix column size}
  \begin{tabular}{|p{3cm}|*{8}{r|}|} \hline
%     &&&&&& \\
    Model / Domain & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{\domain{avg}} & \multicolumn{1}{c||}{w\domain{avg}} \\ \hline % & \multicolumn{1}{c|}{\domain{news}} 
  %    \system{Mixed-Nat}  & 37.3 & 54.6 & 50.1 & 33.5 & 43.2 & 77.5  & 41.1  & 49.4 \\% & 23.5\\
  %     \system{FT-Full}       & 37.7 & \SB{59.2} & \SB{54.5} & 34.0 & \SB{46.8} & \SB{90.8}   & 42.8 & 53.8\\
  %   Full-finetuned on extended in-domain corpora (news) & && 33.96&&& & &\\nn
    \system{DC-Tag}       & 38.5 & \SW{54.0} & 49.0   & 33.6 & \SW{42.2} & \SW{76.7} & 49.0 & 41.6 \\%    & 21.8 \\
    \system{DC-Feat}      & 37.3  & 54.2 & 49.3   & 33.6 & \SW{41.9} & \SW{75.8} & 48.3 & 40.7  \\% & \SW{21.7} \\
    \system{LDC-Tag}     & 37.4   & 54.1 & \SW{48.7} & \SW{32.5} & \SW{41.4} & \SW{75.9} & 48.3 & 39.1         \\% & 22.1 \\ 
    \system{TTM}            & 37.4 & \SW{53.7} & 48.9 & 32.8 & 41.3 & \SW{75.8} & 48.3 & 40.7   \\% &  23.4 \\
    \system{DM}             & 35.4 & 49.3  & 45.2 & 29.7 & 37.1 & \SW{60.0} & 42.8 & 37.8 \\ % & 22.6\\
    \system{ADM}           & 36.1 & 53.5  & 48.0 & 32.0 & 41.1 & 72.1 & 47.1 & 39.5\\% & 23.3 \\
    \system{WDCMT}       & 36.0 & 53.8 & 48.5 & 31.0 & 39.0 & 63.5 & 45.3 & 39.1\\ % & 20.4 \\
    \system{MDL Res}     & 37.3 & 55.5  & \SW{50.2}   & \SW{32.2}   &  \SW{42.1}  & \SW{86.7} & 50.7& 41.2\\%  & \SW{21.2} \\
     \hline
  \end{tabular}
  \caption{Translation performance with automatic domains, computed with the original test sets.}
  \label{tab:subdomains}
  \fyDone{Fill the table with correct results,}
\end{table*}

Table~\ref{tab:subdomains} reports the effect of using automatic domain tags on the 6 domain-specific test sets: each sentence was first assigned to an automatic class, translated with the corresponding multi-domain system; aggregate numbers were then computed, as well as comparisons with the 6-domain scenarios. The results are very clear and confirm previous observations: even though some clusters are very close, the net effect is a decrease in performance for almost all systems and conditions. In this setting, the best MDMT of our pool (MDL-Res) is no longer able to surpass the \system{Mix-Nat} baseline.\fyTodo{Note about scaling?}\fyTodo{More comparison?}
